Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"

### 2020-08-10 ehp_revision/response_to_reviewers.Rmd

The dataset provides exposure by group/area (county), not by individual. These
ecological exposure estimates require us to assume that the exposure level is
the same for all individuals in that county. It ignores/smooths over
within-county variability in exposure.

For wind and rain, we're essentially basing exposure assessment on modeling of
exposure. For rain, this is a based on a re-analysis model that
incorporates/integrates observed data from the area and time, using a model to
integrate data from diverse sources. For wind, the exposure estimates are based
on a model, with a core based on the Willoughby wind model. For flooding and
tornadoes, the exposure is based on records of events anywhere in the county,
and in many cases may be based, essentially, on exposure assessment, as in many
cases someone like someone at a NOAA [?] center have determined whether the
event should be reported in the database.

Even exposure at each person's residence would not suffice, since people move 
around during the day (although perhaps less for most people during a severe
storm).

Data that are aggregated across a population---for example, the total number of
deaths in a geographic area in a certain time period---is known as *ecological
data*, and studies that base inference on this type of data are known as
*ecological studies*. Terms include *ecological*, *aggregate*, *contextual
level*.

### 2020-08-15 ehp_revision/response_to_reviewers.Rmd

**Response:**

\begin{shaded}
\textbf{Reviewer 1 (R1) Comment 1:}
I very much appreciate the rationale for authors' decision to focus at the
county level, especially given the available health-related data. But, since
tropical cyclones don't necessarily have consistent impacts across an entire
county, I think the limitations of this decision are worthy of further
discussion in the discussion section.
\end{shaded}

**Response:**

This is an excellent point. We have add text to the Discussion:

The dataset we present here provides exposure assessments at the county level.
This spatial scale allows for easy integration with health outcome data
aggregated at the county level. Such aggregated or ecological data is often used
for disaster epidemiology, as aggregated data may be easier to access than
individual-level data, especially at a scale that covers many locations and a
long time period and so allows higher statistical power and a broader range of
exposure levels within the study [@wakefield2008overcoming].

These ecological exposure estimates, however, set a common exposure level
throughout the county, ignoring within-county variability, even though such
variability often exists. For some hazards, this within-county variation in
exposure level could be stark. For example, tornadoes cause very localized
damage, directly along the tornado's path---a tornado can destroy homes on one
side of a street while leaving those on the opposite side untouched. Levels of
other hazards, like storm-associated winds and rain, will also vary within a
county, but it will be unlikely that a county will have one area that is exposed
to extremes of these hazards from a tropical cyclone while other parts of the
county are completely unexposed, as both the wind fields and rain fields of
tropical cyclones tend to be large in comparison to the size of a county.

Aggregated data can be used to infer contextual level associations---for
example, the association between county-level exposure to a storm hazard and
county-wide rates of a health outcome. However, ecological data is also
sometimes used to infer individual-level associations (e.g., the association
between personal exposure to a storm hazard and personal risk of  experiencing
the health outcome). Individual-level inference from ecological/aggregated data
is prone to ecological bias [].

-----------------------------------------------------------------------------

### 2020-08-15 ehp_revision/response_to_reviewers.Rmd

**Response:**

\begin{shaded}
\textbf{Reviewer 1 (R1) Comment 1:}
I very much appreciate the rationale for authors' decision to focus at the
county level, especially given the available health-related data. But, since
tropical cyclones don't necessarily have consistent impacts across an entire
county, I think the limitations of this decision are worthy of further
discussion in the discussion section.
\end{shaded}

**Response:**

This is an excellent point. We have add text to the Discussion on the potential
limitations of using aggregated, county-level exposure data in a research study:

> **Text added to Discussion (added text in bold):**
**"The dataset we present here provides exposure assessments at the county level.
This spatial scale allows for easy integration with health outcome data
aggregated at the county level. Such aggregated or ecological data is often used
for disaster epidemiology, as aggregated data may be easier to access than
individual-level data, especially at a scale that covers many locations and a
long time period and so allows higher statistical power and a broader range of
exposure levels within the study [@wakefield2008overcoming].**

> "**These ecological exposure estimates, however, set a common exposure level
throughout the county, ignoring within-county variability, even though such
variability often exists. For some hazards, this within-county variation in
exposure level could be stark. For example, tornadoes cause very localized
damage, directly along the tornado's path---a tornado can destroy homes on one
side of a street while leaving those on the opposite side untouched. Levels of
other hazards, like storm-associated winds and rain, will also vary within a
county, but typically with smoother variation. In particular, it will be
unlikely that a county will have one area that is exposed to extremes of these
hazards from a tropical cyclone while other parts of the county are completely
unexposed, as both the wind fields and rain fields of tropical cyclones tend to
be large in comparison to the size of a county.**

> "**Aggregated data can be used to infer contextual level associations---for
example, the association between county-level exposure to a storm hazard and
county-wide rates of a health outcome. However, ecological data is also
sometimes used to infer individual-level associations (e.g., the association
between personal exposure to a storm hazard and personal risk of  experiencing
the health outcome). Individual-level inference from ecological/aggregated data
is susceptible to ecological bias [@greenland1994invited; @portnov2007ecological;
@idrovo2011three]. Researchers who use the data provided here for ecological
studies---with the aim of making individual-level inferences---should be aware
of this potential and could explore approaches for minimizing risk of ecological
bias (e.g., @wakefield2008overcoming).**"

-----------------------------------------------------------------------------

### 2020-08-15 ehp_revision/writing/ehp_discussion.tex
Work some on revisions

Second, these data are aggregated to the county level. Many health outcome
datasets are aggregated at this level, but some may be aggregated at a finer
spatial resolution (e.g., census tract or ZIP code) or unaggregated (i.e.,
point locations for each outcome). We have published the wind model used to
create this dataset as its own open-source R package
\parencite{stormwindmodel}, and it can be used to model storm-associated winds
at a finer spatial resolution; however, measurements of other hazards cannot
similarly be re-scaled through tools we provide.

>>>>

Second, these data are aggregated to the county level. \textbf{This spatial
scale allows for easy integration with health outcome data aggregated at the
county level. Such aggregated or ecological data is often used for disaster
epidemiology, as aggregated data may be easier to access than individual-level
data, especially at a scale that covers many locations and a long time period
and so allows higher statistical power and a broader range of exposure levels
within the study \parencite{wakefield2008overcoming}.}

\textbf{These ecological exposure estimates, however, set a common exposure
level throughout the county, ignoring within-county variability, even though
such variability often exists. For some hazards, this within-county variation
in exposure level could be stark. For example, tornadoes cause very localized
damage, directly along the tornado's path---a tornado can destroy homes on one
side of a street while leaving those on the opposite side untouched. Levels of
other hazards, like storm-associated winds and rain, will also vary within a
county, but typically with smoother variation. In particular, it will be
unlikely that a county will have one area that is exposed to extremes of these
hazards from a tropical cyclone while other parts of the county are completely
unexposed, as both the wind fields and rain fields of tropical cyclones tend to
be large in comparison to the size of a county.}

\textbf{Aggregated data can be used to infer contextual level
associations---for example, the association between county-level exposure to a
storm hazard and county-wide rates of a health outcome. However, ecological
data is also sometimes used to infer individual-level associations (e.g., the
association between personal exposure to a storm hazard and personal risk of
experiencing the health outcome). Individual-level inference from
ecological/aggregated data is susceptible to ecological bias
\parencite{greenland1994invited, portnov2007ecological, idrovo2011three}.
Researchers who use the data provided here for ecological studies---with the
aim of making individual-level inferences---should be aware of this potential
and could explore approaches for minimizing risk of ecological bias (e.g.,
\cite{wakefield2008overcoming}).}

Further, while many health outcome datasets are aggregated by county, 
some may be aggregated at a finer spatial resolution (e.g., census tract or ZIP
code) or unaggregated (i.e., point locations for each outcome). We have
published the wind model used to create this dataset as its own open-source R
package \parencite{stormwindmodel}, and it can be used to model
storm-associated winds at a finer spatial resolution; however, measurements of
other hazards cannot similarly be re-scaled through tools we provide.
